---
title: "31-alpha_diversity"
title-block-banner: true
author:
  - name: Jinling Xue
    orcid: 0000-0002-8191-4140
toc: true
code-fold: true
format: 
  html:
    embed-resources: true
    smooth-scroll: true
params:
  name: "31-alpha_diversity"
---

Updated: `r format(Sys.time(), '%Y-%m-%d %H:%M:%S', tz = 'CET')` CET.


```{r setup}
#| message: false
#| include: false
here::i_am(paste0(params$name, ".qmd"), uuid = "53b30f44-6e4d-4047-9c49-9d2246519948")
projthis::proj_create_dir_target(params$name, clean = FALSE)
path_target <- projthis::proj_path_target(params$name)
path_source <- projthis::proj_path_source(params$name)
path_data <- path_source("00-raw", params$name)
dir.create(path_data, recursive = T)
```

The purpose of this document is ...

```{r packages}
#| message: false
#| include: false
library(here)
library(tidyverse)
library(vegan)
library(picante)
library(ggpubr)
```

## Tasks

The first task is ...

### Calculate alpha diversity

Code from: `~/project/p0064gvhd/workflow/data/00-raw/result/Alpha Diversity/Alpha-Calculate.R`

```{r}
#| message: false
alpha <- function(x, tree = NULL, base = exp(1)) {
  est <- estimateR(x)
  Richness <- est[1, ]
  Chao1 <- est[2, ]
  ACE <- est[4, ]
  Shannon <- diversity(x, index = 'shannon', base = base)
  Simpson <- diversity(x, index = 'simpson')    #Gini-Simpson ??????
  InvSimpson <- diversity(x, index = 'invsimpson') 
  Pielou <- Shannon / log(Richness, base)
  goods_coverage <- 1 - rowSums(x == 1) / rowSums(x)
  
  result <- data.frame(Richness, Shannon, Simpson, InvSimpson,Pielou, Chao1, ACE, goods_coverage)
  if (!is.null(tree)) {
    PD_whole_tree <- pd(x, tree, include.root = FALSE)[1]
    names(PD_whole_tree) <- 'PD_whole_tree'
    result <- cbind(result, PD_whole_tree)
  }
  result
}

fin_alpha <- here(path_data, 'Alpha_AF_2.txt')
fout_alpha <- path_target('Alpha_AF_Final_2.csv')

otu <- read.delim(fin_alpha, row.names = 1, sep = '\t', stringsAsFactors = FALSE, check.names = FALSE)
otu <- t(otu)
alpha_all <- alpha(otu, base = 2)
write.csv(alpha_all, fout_alpha, quote = FALSE)
```

### Inverse simpson ABX

Code from: `~/project/p0064gvhd/workflow/data/00-raw/result/Alpha Diversity/Inverse-simpson-ABX.R`

```{r}
#| message: false
fin_abx <- here(path_data, 'Alpha_ABX_Final.txt')
fout_abx <- path_target("Alpha-InvSimpson-5D-Tall.png")
otu <- read.delim(fin_abx, sep = '\t', stringsAsFactors = FALSE, check.names = FALSE)
# theme_bw() +
#   theme(axis.line = element_line(color='black'),
#         plot.background = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.border = element_blank())

otu$Group<-factor(otu$Group,levels = c("No ABX","ABX"))
my_comparisons <- list(c("No ABX", "ABX"))

p0<-ggboxplot(otu,
              x="Group",
              y="InvSimpson",
              color="Group",
              palette = "jama",
              add = "jitter")
p0
p1<-p0+color_palette(c("#0F7FFE", "#FB0106"))+labs( y = "Inverse Simpson",x="")+theme(legend.position = "none", axis.text.x=element_text(family="Arial",colour="black",angle = 45,vjust = 0.15,hjust = 0.5))
p1
p2<-p1+stat_compare_means(comparisons=my_comparisons,label = "p.signif", label.x = 1.5, label.y = 25)
p2
p3<- ggpar(p2, ylim = c(0, 30))
p3
png(file = fout_abx, width =700,height = 1400,units = "px",res =300) 
dev.off()
```

### Inverse simplson GVHD

code from: `~/project/p0064gvhd/workflow/data/00-raw/result/Alpha Diversity/Inverse-simpson-GVHD.R`

```{r}
#| message: false
# fin_1e <- here(path_data, 'Alpha-1E-Final.txt')
fin_1e <- here(path_data, 'Alpha_GVHD_Final.txt')
fout_1e <- path_target("Alpha-Inversesimpson-5a.png")

otu <- read.delim(fin_1e, sep = '\t', stringsAsFactors = FALSE, check.names = FALSE)
# theme_bw() +
#   theme(axis.line = element_line(color='black'),
#         plot.background = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.border = element_blank())

otu$Group<-factor(otu$Group,levels = c("No GI-GvHD","GI-GvHD"))

p0<-ggboxplot(otu,
              x="Group",
              # y="Inversesimpson",
              y="InvSimpson",
              color="Group",
              palette = "jama",
              add = "jitter",
              width =0.5)
p0
p1<-p0+color_palette(c("#0F7FFE", "#FB0106"))+labs( y = "Inverse Simpson",x="")+theme(legend.position = "none")
p1
p2<-p1+stat_compare_means(aes(label = ..p.signif..),
                          method = "t.test", ref.group = "No GI-GvHD")+theme(axis.text.x=element_text(family="Arial",colour="black",angle = 45,vjust = 0.15,hjust = 0.5))
p2
p3<- ggpar(p2, ylim = c(0, 30))
p3
png(file = fout_1e, width =1800,height = 1400,units = "px",res =300) 
dev.off()
```


### Alpha-1S

code from: `~/project/p0064gvhd/workflow/data/00-raw/result/Alpha Diversity/Alpha-1S.R`

```{r}
fin_1s <- here(path_data, 'Alpha-1S-Final.txt')
fout_1s <- path_target("Alpha-Shannon-1s.png")

otu <- read.delim(fin_1s, sep = '\t', stringsAsFactors = FALSE, check.names = FALSE)
# theme_bw() +
#   theme(axis.line = element_line(color='black'),
#         plot.background = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.border = element_blank())

otu$Group<-factor(otu$Group,levels = c("Donor","MUC","RGB"))
p0<-ggboxplot(otu,
              x="Group",
              y="Shannon",
              color="Group",
              palette = "jama",
              add = "jitter")
p0
p1<-p0+color_palette(c("#CCCCCB", "#107F01","#FC8008"))+labs( y = "Effective Shannon",x="")+theme(legend.position = "right")
p1
p2<-p1+theme(text=element_text(family="Arial",colour="black"))
p2
p3<- ggpar(p2, ylim = c(0, 600))
p3
png(file = fout_1s, width =1800,height = 1400,units = "px",res =300) 
print(p3)
dev.off()
```


### Alpha-4A

code from: `~/project/p0064gvhd/workflow/data/00-raw/result/Alpha Diversity/Alpha-4A.R`

```{r}
fin_4e <- here(path_data, 'Alpha-4E-Final.txt')
fout_4e <- path_target("Alpha-Shannon-4Anew.png")

otu <- read.delim(fin_4e, sep = '\t', stringsAsFactors = FALSE, check.names = FALSE)
# theme_bw() +
#   theme(axis.line = element_line(color='black'),
#         plot.background = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.border = element_blank())

otu$Group<-factor(otu$Group,levels = c("No ABX","ABX"))
my_comparisons <- list(c("No ABX", "ABX"))
p0<-ggboxplot(otu,
              x="Group",
              y="Shannon",
              color="Group",
              palette = "jama",
              add = "jitter")
p0
p1<-p0+color_palette(c("#0F7FFE", "#FB0106"))+labs( y = "Effection Shannon",x="")+theme(legend.position = "none", axis.text.x=element_text(family="Arial",colour="black",angle = 45,vjust = 0.15,hjust = 0.5))
p1
p2<-p1+stat_compare_means(comparisons=my_comparisons)
p2
p3<- ggpar(p2, ylim = c(0, 1000))
p3
png(file = fout_4e, width =1800,height = 1400,units = "px",res =300) 
print(p3)
dev.off()
```


### Alpha-Antifungals

code from: `~/project/p0064gvhd/workflow/data/00-raw/result/Alpha Diversity/Alpha-Antifungals.R`

```{r}
fin_gvhd <- here(path_data, 'Alpha_GVHD_Final.txt')
fout_gvhd <- path_target("Alpha-Richness-5A-TALL.png")

otu <- read.delim(fin_gvhd, sep = '\t', stringsAsFactors = FALSE, check.names = FALSE)
# theme_bw() +
#   theme(axis.line = element_line(color='black'),
#         plot.background = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.border = element_blank())

otu$Group<-factor(otu$Group,levels = c("No GI-GvHD","GI-GvHD"))
my_comparisons <- list(c("No GI-GvHD", "GI-GvHD"))
p0 <- ggboxplot(otu,
              x="Group",
              y="Richness",
              color="Group",
              palette = "jama",
              add = "jitter",
              width =0.5)
p1<-p0+color_palette(c("#0F7FFE", "#FB0106"))+labs( y = "Richness",x="")+theme(legend.position = "none")

p2<-p1+stat_compare_means(comparisons=my_comparisons,label = "p.signif", label.x = 1.5, label.y = 2000)+theme(axis.text.x=element_text(family="Arial",colour="black",angle = 45,vjust = 0.15,hjust = 0.5))
p2
p3<- ggpar(p2, ylim = c(0, 2500))

ggsave(fout_gvhd, p3, width =700,height = 1400,units = "px" )
#png(file = "Alpha-Richness-5A-TALL.png",width =700,height = 1400,units = "px",res =300) 
```


### Alpha-GvHD-1A

code from: `~/project/p0064gvhd/workflow/data/00-raw/result/Alpha Diversity/Alpha-GvHD-1A.R`

```{r}
fin_1a <- here(path_data, "Alpha-1A.txt")
fout_1a <- path_target("Alpha-Shannon-1Anew.png")

otu <- read.delim(fin_1a, sep = '\t', stringsAsFactors = FALSE, check.names = FALSE)
# theme_bw() +
#   theme(axis.line = element_line(color='black'),
#         plot.background = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.border = element_blank())

otu$Timepoints<-factor(otu$Timepoints,levels = c("Day -7","Day 0","Day +7","Day +14","Day +21","Day +28","> Day +35"))
my_comparisons <- list(c("Day -7", "Day +14")) 
p0<-ggboxplot(otu,
             x="Timepoints",
             y="Shannon",
             color="Timepoints",
             palette = "jama",
             add = "jitter")
p0
p1<-p0+color_palette(c("#3F7F02", "#FDCC65", "#FC8008","#FC6665","#FB0106","#FB027F","#800080"))+labs( y = "Richness",x="")+theme(legend.position = "right")
p1
p2<-p1+theme(axis.text.x=element_text(family="Arial",colour="black",angle = 45, vjust = 0.3,hjust = 0.5))
p2
p3<-p2+stat_compare_means(comparisons=my_comparisons)
p3
p4<- ggpar(p3, ylim = c(0, 1000))
png(file = fout_1a, width =1800,height = 1400,units = "px",res =300) 
print(p4)
dev.off()
```


### Alpha-Richness-4E

code from: `~/project/p0064gvhd/workflow/data/00-raw/result/Alpha Diversity/Alpha-Richness-4E.R`

```{r}
fin_early <- here(path_data, "Alpha-Early-final.txt")
fout_early <- path_target("Alpha-Shannon-ABX.png")
otu <- read.delim(fin_early, sep = '\t', stringsAsFactors = FALSE, check.names = FALSE)

# theme_bw() +
#   theme(axis.line = element_line(color='black'),
#         plot.background = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.border = element_blank())

otu$Group<-factor(otu$Group,levels = c("No ABX","Early ABX","Late ABX"))
my_comparisons <- list(c("No ABX","Early ABX"))
p0<-ggboxplot(otu,
              x="Group",
              y="Shannon",
              color="Group",
              palette = "jama",
              add = "jitter")
p0
p1<-p0+color_palette(c("#90EE90", "#FFC0CB","#DC143C"))+labs( y = "Shannon effective",x="")+theme(legend.position = "none",text=element_text(family="Arial",colour="black"))
p1
p2<-p1+stat_compare_means(label.y = 750)+stat_compare_means(comparisons=my_comparisons)
p2
p3<- ggpar(p2, ylim = c(0, 1000))
p3
png(file = fout_early, width =1800,height = 1400,units = "px",res =300) 
dev.off()
```

### Dot Alpha

code from: `~/project/p0064gvhd/workflow/data/00-raw/result/Alpha Diversity/Dotsplot-Alpha.R`

```{r}
fin_time <- here(path_data, "Alpha_TIME_Final.txt")
fout_time <- path_target("Alpha-InvSimpson-Timenew.png")

otu <- read.delim(fin_time, sep = '\t', stringsAsFactors = FALSE, check.names = FALSE)

otu$Timepoints<-factor(otu$Timepoints,levels = c("Day -7","Day 0","Day +7","Day +14","Day +21","Day +28"))

otu <- otu %>%
  mutate(timepoints2=str_replace(Timepoints, "^.*Day ", "")) %>%
  mutate(timepoints2 = as.integer(timepoints2)) %>%
  mutate(timepoints3 = as.character(timepoints2))
unique(otu$timepoints2)
p0 <- ggplot(otu,aes(x=timepoints2,y=InvSimpson))+
  #geom_point(aes(color=Center))+
  geom_point(position=position_jitter(h=2, w=2),alpha = 0.5, size = 1)+
  geom_smooth(aes(color=Center))+
  geom_smooth(color="black")+
  scale_x_continuous(breaks= otu$timepoints2,
                   labels=otu$Timepoints)+
  theme_bw()+
  theme(legend.position = "right",panel.grid.major = element_blank(),panel.grid.minor = element_blank())
p0
p1<-p0+color_palette(c("#FFB562", "#F87474"))+labs( y = "Inverse Simpson",x="")
p1
p2<-p1+
  theme(axis.text.x=element_text(family="Arial",colour="black",angle = 45, vjust = 0.3,hjust = 0.5),axis.text.y=element_text(family="Arial",colour="black"))
p2
p3<- ggpar(p2, ylim = c(0, 30))
png(file = fout_time, width =1200,height = 800,units = "px",res =300) 
print(p3)
dev.off()

```


### Inverse-simpson-GVHD

code from: `~/project/p0064gvhd/workflow/data/00-raw/result/Alpha Diversity/Inverse-simpson-GVHD.R`

```{r}
fin_inv <- here(path_data, "Alpha-1E-Final.txt")
fout_inv <- path_target("Alpha-Inversesimpson-5a_tmp2.png")

otu <- read.delim(fin_inv, sep = '\t', stringsAsFactors = FALSE, check.names = FALSE)
# theme_bw() +
#   theme(axis.line = element_line(color='black'),
#         plot.background = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.border = element_blank())

otu$Group<-factor(otu$Group,levels = c("No GI-GvHD","GI-GvHD"))

p0<-ggboxplot(otu,
              x="Group",
              y="Inversesimpson",
              color="Group",
              palette = "jama",
              add = "jitter",
              width =0.5)
p0
p1<-p0+color_palette(c("#0F7FFE", "#FB0106"))+labs( y = "Inverse Simpson",x="")+theme(legend.position = "none")
p1
p2<-p1+stat_compare_means(aes(label = ..p.signif..),
                          method = "t.test", ref.group = "No GI-GvHD")+theme(axis.text.x=element_text(family="Arial",colour="black",angle = 45,vjust = 0.15,hjust = 0.5))
p2
p3<- ggpar(p2, ylim = c(0, 125))
p3
png(file = fout_inv, width =1800,height = 1400,units = "px",res =300) 
dev.off()
```





## Files written

These files have been written to the target directory, ```r paste0("data/", params$name)```:

```{r list-files-target}
projthis::proj_dir_info(path_target())
```
